---
title: ""
format: html
editor: visual
execute: 
  echo: false
---

# Retail Growth Opportunities Across the United States (2019–2022)

## Executive Summary

This report analyses retail sector performance across U.S. states using year-on-year (YoY) sales growth data between 2019 and 2022. The objective is to identify sectors that demonstrate strong growth trends and highlight geographical differences in retail performance.

The analysis found that Miscellaneous Store Retailers recorded the highest average YoY growth across the period, followed by Sporting Goods, Hobby, Musical Instrument and Book Stores. Food and Beverage Retailers demonstrated lower growth rates but appear to offer greater long-term stability due to consistent consumer demand.

For entrepreneurs, high-growth sectors may represent attractive opportunities for market entry. For investors, balancing growth potential against business stability remains critical when evaluating retail opportunities.

### Key Findings

-   Miscellaneous Store Retailers achieved the highest average YoY growth.
-   Electronics and Appliance Retailers showed the weakest average growth.
-   Retail performance varied significantly between states.
-   Growth rates alone should not be used as the sole measure of business attractiveness.

------------------------------------------------------------------------

# Business Problem

Choosing the right retail sector is one of the most important decisions for entrepreneurs and investors. Entering a sector with strong market growth can increase the likelihood of revenue expansion, while selecting a poor-performing sector may limit growth opportunities.

This analysis seeks to answer two business questions:

1.  Which retail sub-sectors experienced the strongest growth between 2019 and 2022?
2.  Which retail sub-sectors performed best within each U.S. state?

The findings provide insight into potential market opportunities and can support business planning, investment decisions and market research activities.

------------------------------------------------------------------------

# Data Source

The analysis uses the State Retail Sales data-set from the 2022 Tidy-Tuesday project.

The data-set contains:

-   U.S. state retail sales information
-   Retail sub-sector classifications
-   Year-on-year sales growth percentages
-   Monthly observations across multiple years

For this analysis:

-   National totals were removed.
-   Aggregate "total" sub-sectors were excluded.
-   Year-on-year growth percentages were converted into numeric values.
-   State-level observations were analysed.

------------------------------------------------------------------------

# Methodology

The analysis was completed in three stages.

### Stage 1: Sector Performance Analysis

Average YoY growth was calculated for each retail sub-sector across all states.

This allows identification of sectors that consistently achieved stronger growth over the analysis period.

### Stage 2: State-Level Opportunity Analysis

For each state, the retail sub-sector with the highest average YoY growth was identified.

This helps reveal whether certain sectors perform consistently well across multiple geographic markets.

### Stage 3: Variability Assessment

Standard errors were calculated for each sub-sector to assess variation in performance and identify sectors with more volatile growth patterns.

------------------------------------------------------------------------

# Findings

```{r echo=FALSE}
#| warning: false
#| message: false

state_retail <- readr::read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-13/state_retail.csv"
)
library(tidyverse)
library(readr)

# -----------------------
# CLEAN DATA
# -----------------------
df_all <- state_retail %>%
  mutate(change_yoy = parse_number(change_yoy)) %>%
  filter(!is.na(change_yoy)) %>%
  filter(state_abbr != "USA") %>%
  
  # REMOVE total rows
  filter(!str_detect(str_to_lower(subsector), "total"))

# -----------------------
# SUMMARISE ACROSS STATES
# -----------------------
df_subsector_summary <- df_all %>%
  group_by(subsector) %>%
  summarise(
    avg_change = mean(change_yoy, na.rm = TRUE),
    se_change = sd(change_yoy, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_change))

# -----------------------
# PLOT
# -----------------------
ggplot(
  df_subsector_summary,
  aes(
    x = reorder(subsector, avg_change),
    y = avg_change
  )
) +
  
  geom_col(fill = "steelblue") +
  
  geom_errorbar(
    aes(
      ymin = avg_change - se_change,
      ymax = avg_change + se_change
    ),
    width = 0.2,
    color = "darkred"
  ) +
  
  coord_flip() +
  
  labs(
    title = "Average YoY Change by Subsector",
    x = "Subsector",
    y = "Average YoY Change"
  ) +
  
  theme_minimal()

```

```{r}
#| message: false
#| warning: false
#| results: asis
#| fig-width: 14
#| fig-height4: 5
library(tidyverse)
library(readr)
library(ggforce)

# -----------------------
# LOAD & CLEAN DATA
# -----------------------
state_retail <- readr::read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-13/state_retail.csv"
)

df <- state_retail %>%
  mutate(change_yoy = parse_number(change_yoy)) %>%
  filter(!is.na(change_yoy)) %>%
  filter(state_abbr != "USA") %>%
  filter(!str_detect(str_to_lower(subsector), "total"))

# -----------------------
# FIND BEST SUBSECTOR
# -----------------------
best_subsector <- df %>%
  group_by(state_abbr, subsector) %>%
  summarise(mean_change = mean(change_yoy, na.rm = TRUE), .groups = "drop") %>%
  group_by(state_abbr) %>%
  slice_max(mean_change, n = 1, with_ties = FALSE) %>%
  ungroup()

# -----------------------
# FILTER TO BEST SUBSECTOR ONLY
# -----------------------
df_best <- df %>%
  inner_join(best_subsector, by = c("state_abbr", "subsector"))

# -----------------------
# SUMMARISE BY YEAR + SE
# -----------------------
df_summary <- df_best %>%
  group_by(state_abbr, subsector, year) %>%
  summarise(
    avg_change = mean(change_yoy, na.rm = TRUE),
    se_change = sd(change_yoy, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# -----------------------
# AUTOMATIC PAGE COUNT
# -----------------------
n_states <- length(unique(df_summary$state_abbr))
pages <- ceiling(n_states / 5)

#| results: asis
#| fig-width: 14
#| fig-height: 4
#| warning: false

# -----------------------
# LOOP OVER ALL PAGES TO GENERATE TABS
# -----------------------
for (i in seq_len(pages)) {
  
  # 1. Print the markdown heading for the tab
  cat(paste0("\n\n### Page ", i, "\n\n"))
  
  # 2. Build the plot for this specific page
  p <- ggplot(df_summary, aes(x = year, y = avg_change)) +
    geom_col(fill = "steelblue") +
    geom_errorbar(
      aes(
        ymin = avg_change - se_change,
        ymax = avg_change + se_change
      ),
      width = 0.2,
      color = "darkred"
    ) +
    facet_wrap_paginate(
      ~ paste(state_abbr, "\n", subsector),
      ncol = 5,
      nrow = 1,
      page = i,
      scales = "free_y"
    ) +
    labs(
      title = paste("Best Subsector by State — Page", i),
      x = "Year",
      y = "YoY Change"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # 3. Print the plot object so Quarto renders it inside the tab
  print(p)
}
```

## Finding 1: Miscellaneous Store Retailers Recorded the Highest Growth

The first visualisation shows average YoY growth for each retail subsector.

Miscellaneous Store Retailers recorded the highest average growth rate across the dataset.

This suggests that businesses operating within this category experienced stronger relative expansion compared with most other retail sectors.

However, it is important to note that higher growth rates do not necessarily imply higher revenue or profitability. Smaller sectors often achieve larger percentage increases due to lower starting sales volumes.

### Business Implication

Entrepreneurs may wish to investigate niche retail opportunities within this category, particularly in markets demonstrating strong local demand.

------------------------------------------------------------------------

## Finding 2: Sporting Goods and Hobby Retailers Performed Strongly

Sporting Goods, Hobby, Musical Instrument and Book Retailers also displayed above-average growth.

This may indicate continued consumer spending on recreational activities, hobbies and personal interests during the period analysed.

### Business Implication

Businesses operating within enthusiast-driven markets may benefit from strong customer engagement and repeat purchases.

------------------------------------------------------------------------

## Finding 3: Electronics and Appliance Retailers Showed Lower Growth

Electronics and Appliance Retailers recorded some of the lowest average YoY growth rates.

While these businesses may generate substantial revenue, growth appears slower relative to other retail sectors.

### Business Implication

Businesses entering this sector may face mature market conditions, increased competition and slower expansion opportunities.

------------------------------------------------------------------------

## Finding 4: State-Level Differences Exist

The state-level analysis demonstrated that the highest-performing sub-sector varies across states.

Although Miscellaneous Store Retailers frequently appeared among top-performing categories, no single sub-sector dominated every market.

### Business Implication

Location remains a critical factor in retail success. Businesses should conduct local market research before selecting a retail category or expansion strategy.

------------------------------------------------------------------------

# Strategic Recommendations

## Recommendation 1: Prioritise Market Research Over National Trends

National averages provide useful indicators, but local market conditions should drive decision-making.

Before entering any sector, businesses should assess:

-   Local competition
-   Consumer demographics
-   Population growth
-   Disposable income levels
-   Commercial property costs

------------------------------------------------------------------------

## Recommendation 2: Consider High-Growth Sectors

Entrepreneurs seeking growth opportunities should further investigate:

-   Miscellaneous Store Retailers
-   Sporting Goods Retailers
-   Hobby and Recreational Retailers

These sectors demonstrated the strongest average growth within the dataset.

------------------------------------------------------------------------

## Recommendation 3: Balance Growth and Stability

High growth does not always equal lower risk.

Food and Beverage Retailers demonstrated lower growth rates but operate within a market supported by consistent consumer demand.

Investors may prefer sectors that balance moderate growth with predictable revenue streams.

------------------------------------------------------------------------

# Limitations

Several limitations should be considered when interpreting these findings.

-   The data-set measures growth rather than profitability.
-   Business costs are not included.
-   Consumer demographic data is not included.
-   E-commerce performance is not analysed separately.
-   COVID-19 likely influenced retail trends during the study period.
-   State-level averages may hide local differences between cities and regions.

Future analysis could combine retail growth data with demographic, income and population statistics to generate more actionable business insights.

------------------------------------------------------------------------

# Conclusion

This analysis identified significant differences in retail sector performance across the United States between 2019 and 2022.

Miscellaneous Store Retailers recorded the highest average YoY growth, while Sporting Goods and Hobby Retailers also demonstrated strong performance. However, growth rates alone do not determine business success.

The most effective business decisions combine market growth data with local demand, competitive analysis and financial considerations.

For entrepreneurs and investors, the findings provide a useful starting point for identifying promising retail opportunities and conducting deeper market research.
